---
title: "RTKpin_boxplots"
author: "Becca Morris"
date: "2024-05-02"
output: html_document
---

```{r setup, include=FALSE}
library(readxl)
library(tidyverse)
library(dplyr)
library(lubridate)
```
#1. Run these:
```{r}
setwd("~/GitHub/BeccaTransects/data")

RTKpindata_norefs <- read_xlsx("RTKSedimentPinData.xlsx", sheet = "RTKData_thesispins_norefs_pivot")

```

#prepping the data frame:
##*remember that this df is still in ft NADV88 and isn't converted until the next step
```{r}
RTKpins <- RTKpindata_norefs %>% 
  select(-c("midg","midt","upt","upg", "lowt", "lowg", "Date")) %>% #can remove these columns bc the top-ground calculations are already calculated in excel in their own columns
  mutate_at(c('upper', 'middle'), as.numeric) %>% #removed the POSITX (lubridate) version of date
  mutate(Date = paste0(Year, "-", Month)) %>% #created my own date to match the sediment pin code
  mutate(across(where(is.numeric), ~round(., 3))) %>%
  group_by(Name, Shoreline_End) %>% 
  rename(Transect_ID = Name)
```

# Setting up df for EvW Plots:
```{r}
RTKpinschangeEvW <- RTKpins %>%
  mutate(UpperDiff = ((last(upper) - first(upper))*-1)*30.48) %>%  # *30.48 converts NADV88 ft to cm
  mutate(MiddleDiff = ((last(middle) - first(middle))*-1)*30.48) %>% 
  mutate(LowerDiff = ((last(lower) - first(lower))*-1)*30.48) %>%
  select(-c("Date", "Year", "Month", "upper", "middle", "lower")) %>% 
  distinct() %>%  #this removes repeated values for each month's calculation
  pivot_longer(UpperDiff:LowerDiff, names_to = "Pin_Location", values_to = "Sediment_Change") %>% 
  rename(Zone = Pin_Location) %>% 
  mutate(Shoreline_End = factor(Shoreline_End, levels = c("west", "east")),
         Zone = factor(Zone, levels = c("UpperDiff", "MiddleDiff", "LowerDiff"))) %>%
  group_by(Shoreline_End, Zone) #%>% 
  #summarize(mean_difference = mean(Sediment_Change, na.rm = TRUE),
  #          sd_difference = sd(Sediment_Change, na.rm = TRUE),
  #          n_difference = n()) %>% 
  #mutate(se_difference = sd_difference/sqrt(n_difference)) %>% 
  #mutate(across(where(is.numeric), ~round(., 1)))
```



# EvW *ONLY* Plot:
```{r}
cbprtk <- c("#009E73", "#56B4E9", "#F0E442" , "#E69F00", "#000000", "#0072B2","#D55E00", "#CC79A7","#999999")
```

#EvW only. No marsh zone or log:
```{r}
RTKpinchangeEvW_box2 <- ggplot(RTKpinschangeEvW,
                                  mapping = aes(x = Shoreline_End,
                                  y = Sediment_Change,
                                  fill = Shoreline_End,
                                  color = Shoreline_End)) +
  geom_boxplot(position = position_dodge(0.55),color = "black", width = 0.50) +
  stat_summary(fun.y="mean", shape = 5, size = 0.4, position = position_dodge(0.55), color = "black") + #this adds the diamond for the means
  geom_text(data = n_values_RTKpins, aes(x = Shoreline_End, y = -25,  # Adjust y-position as needed
                                         label = paste0("n = ", n)),
            position = position_dodge(0.55), vjust = 1, size = 3, color = "black") +
  # stat_summary(
  #   fun = mean,
  #   geom = "text",
  #   aes(label = sprintf("%.2f", after_stat(y))),  # Format mean values
  #   position = position_dodge(0.55),
  #   size = 3,  # Adjust size as needed
  #   vjust = 0.28,# Adjust vertical positioning
  #   hjust = -0.5,
  #   color = "black") +
  geom_hline(yintercept=0, linetype="dashed", color = "black") +
  scale_fill_manual(values = cbprtk) +
  coord_cartesian(ylim = c(-25, 20)) + #sets the limit of the y-axis to show a more dramatic difference bw variables
  labs(
    title = "RTK Measurement: Change in Sediment Pin Length",
    subtitle = "From Jun 2022 to Sep 2023   (Positive value = accretion   Negative value = erosion)",
    y = "Change in Pin Length (cm)",
    x = "End of Shoreline",
    #fill= "End of Shoreline",
    tag = "RTKpin_boxplots.Rmd_RTKpinchangeEvW_box2",
    caption = "*n=10 west/log, n=7 west/no log, n=12 east/log, n=9 east/no log") +
  theme_bw(base_size = 10) +
  theme(axis.title = element_text(face = "bold", size=13),
        legend.title = element_text(face = "bold", size=09),
        legend.text=element_text(size=09),
        legend.position = "none",   #for no legend type: "none"
        legend.background = element_rect(fill="lightgrey",
                                  size=0.5, linetype="solid", 
                                  colour ="darkgrey"),
        axis.text.x = element_text(size = 11),
        axis.text.y = element_text(size = 10),
        plot.title = element_text(face = "bold", size = 14),
        plot.caption = element_text(size = 09),
        plot.subtitle = element_text(size = 11),
        plot.tag = element_text(size = 6, angle = 90),
        plot.tag.position = "right",
        strip.text = element_text(size = 13, color = "black"))
  
  

RTKpinchangeEvW_box2
```


#This calculates the n-value for each shoreline end so that you can add it to the plot below
```{r}
n_values_RTKpins <- RTKpinschangeEvW %>%
  group_by(Zone, Shoreline_End) %>%
  summarise(n = n()) %>%
  ungroup()
```

#This one is with marsh zone included
```{r}
RTKpinchangeEvW_box <- ggplot(RTKpinschangeEvW,
                                  mapping = aes(x = Zone,
                                  y = Sediment_Change,
                                  fill = Shoreline_End,
                                  color = Shoreline_End)) +
  geom_boxplot(position = position_dodge(0.55),color = "black", width = 0.50) +
  stat_summary(fun.y="mean", shape = 5, size = 0.4, position = position_dodge(0.55), color = "black") + #this adds the diamond for the means
  geom_text(data = n_values_RTKpins, 
            aes(x = Zone, y = -25, label = paste0("n = ", n)),
            position = position_dodge(0.55), vjust = 1, size = 3, color = "black") + #this add the n_count to the plot. vjust moves it down, and nudge_x moved it right
  geom_hline(yintercept=0, linetype="dashed", color = "black") +
  scale_fill_manual(values = cbprtk) +
  coord_cartesian(ylim = c(-25, 20)) + #sets the limit of the y-axis to show a more dramatic difference bw variables
  labs(
    title = "RTK Measurement: Change in Sediment Pin Length",
    subtitle = "From Jun 2022 to Sep 2023   (Positive value = accretion   Negative value = erosion)",
    y = "Change in Pin Length (cm)",
    x = "Marsh Zone",
    fill= "End of Shoreline",
    tag = "RTKpin_boxplots.Rmd_RTKpinchangeEvW_box") +
    #caption = "*n=10 west/log, n=7 west/no log, n=12 east/log, n=9 east/no log") +
  theme_bw(base_size = 10) +
  theme(axis.title = element_text(face = "bold", size=13),
        legend.title = element_text(face = "bold", size=09),
        legend.text=element_text(size=09),
        legend.position = "top",   #for no legend type: "none"
        legend.background = element_rect(fill="lightgrey",
                                  size=0.5, linetype="solid", 
                                  colour ="darkgrey"),
        axis.text.x = element_text(size = 11),
        axis.text.y = element_text(size = 10),
        plot.title = element_text(face = "bold", size = 14),
        plot.caption = element_text(size = 09),
        plot.subtitle = element_text(size = 11),
        plot.tag = element_text(size = 6, angle = 90),
        plot.tag.position = "right",
        strip.text = element_text(size = 13, color = "black"))
  
  

RTKpinchangeEvW_box
```



## Setting up df for EvW *AND* Log v No Log:
```{r}
RTKpinchangeLvNL <- RTKpins %>% 
  select(Transect_ID, Shoreline_End, upper, middle, lower, Date, Log_Presence) %>% 
  mutate(UpperDiff = ((last(upper) - first(upper))*-1)*30.48) %>%  # *30.48 converts NADV88 ft to cm
  mutate(MiddleDiff = ((last(middle) - first(middle))*-1)*30.48) %>% 
  mutate(LowerDiff = ((last(lower) - first(lower))*-1)*30.48) %>%
  select(-c('Date', 'upper', 'middle', 'lower')) %>% #remove all unnecessary columns
  distinct() %>%  #this removes repeated values for each month's calculation
  pivot_longer(UpperDiff:LowerDiff, names_to = "Pin_Location", values_to = "Sediment_Change") %>%
  rename(Zone = Pin_Location) %>% 
  mutate(Shoreline_End = factor(Shoreline_End, levels = c("west", "east")),
         Zone = factor(Zone, levels = c("UpperDiff", "MiddleDiff", "LowerDiff"))) %>% #this puts west first in the legend
  group_by(Log_Presence, Shoreline_End, Zone) #%>% 
  #summarize(mean_difference = mean(Sediment_Change, na.rm = TRUE),
  #          sd_difference = sd(Sediment_Change, na.rm = TRUE),
  #          n_difference = n()) %>% 
  #mutate(se_difference = sd_difference/sqrt(n_difference)) %>% 
  #mutate(across(where(is.numeric), ~round(., 1)))
```


```{r}
RTKpinchangeLvNL_box <- ggplot(RTKpinchangeLvNL,
                                  mapping = aes(x = Zone,
                                  y = Sediment_Change,
                                  #group = Log_Presence,
                                  fill = Log_Presence,
                                  color = log_Presence)) +
  geom_boxplot(position = position_dodge(0.55),color = "black", width = 0.50) +
  geom_text(data = n_values_RTKpins_LvNL,                              #adds n-values to plot
            aes(x = Zone, y = -25, label = paste0("n = ", n)),
            position = position_dodge(0.70), vjust = 1, size = 3, color = "black") +
  geom_hline(yintercept=0, linetype="dashed", color = "black") +
  scale_fill_manual(values = cbprtk) +
  facet_grid(~ Shoreline_End) +
  coord_cartesian(ylim = c(-25, 20)) + #sets the limit of the y-axis to show a more dramatic difference bw variables
  scale_x_discrete(labels = c(UpperDiff = "Upper", MiddleDiff = "Middle", LowerDiff = "Lower")) +
  labs(
    title = "RTK Measurement: Change in Sediment Pin Length ",
    subtitle = "From Jun 2022 to Sep 2023   (Positive value = accretion   Negative value = erosion)",
    y = "Change in Pin Length (cm)",
    x = "Marsh Zone",
    fill= "Log Presence",
    tag = "RTKpin_boxplots.Rmd_RTKpinchangeLvNL_box") +
    #caption = "*n=10 west/log, n=7 west/no log, n=12 east/log, n=9 east/no log") +
  theme_bw(base_size = 10) +
  theme(axis.title = element_text(face = "bold", size=13),
        legend.title = element_text(face = "bold", size=09),
        legend.text=element_text(size=09),
        legend.position = "top",   #for no legend type: "none"
        legend.background = element_rect(fill="lightgrey",
                                  size=0.5, linetype="solid", 
                                  colour ="darkgrey"),
        axis.text.x = element_text(size = 11),
        axis.text.y = element_text(size = 10),
        plot.title = element_text(face = "bold", size = 14),
        plot.caption = element_text(size = 09),
        plot.subtitle = element_text(size = 11),
        plot.tag = element_text(size = 6, angle = 90),
        plot.tag.position = "right",
        strip.text = element_text(size = 13, color = "black"))
  
  

RTKpinchangeLvNL_box
#Will say: Warning: Removed 3 rows containing non-finite outside the scale range (`stat_boxplot()`). This is due to having NA's for a few of the pins.  
```

# This counts the observations for each combination
```{r}
n_values_RTKpins_LvNL <- RTKpinchangeLvNL %>%
  group_by(Log_Presence, Shoreline_End, Zone) %>%
  summarise(n = n(), .groups = 'drop')  #drop is the same as 'ungroup'
```


#West vs East with log variable
```{r}
RTKpinchangeEvWandLvNL_box <- ggplot(RTKpinchangeLvNL,
                                  mapping = aes(x = Zone,
                                  y = Sediment_Change,
                                  #group = Log_Presence,
                                  fill = Shoreline_End,
                                  color = Shoreline_End)) +
  geom_boxplot(position = position_dodge(0.55),color = "black", width = 0.50) +
  stat_summary(fun.y="mean", shape = 5, size = 0.4, position = position_dodge(0.55), color = "black") + #this adds the diamond for the means
  geom_text(data = n_values_RTKpins_LvNL,                              #adds n-values to plot
            aes(x = Zone, y = -25, label = paste0("n = ", n)),
            position = position_dodge(0.70), vjust = 1, size = 3, color = "black") +
  geom_hline(yintercept=0, linetype="dashed", color = "black") +
  scale_fill_manual(values = cbprtk) +
  facet_grid(~ Log_Presence) +
  coord_cartesian(ylim = c(-25, 20)) + #sets the limit of the y-axis to show a more dramatic difference bw variables
  scale_x_discrete(labels = c(UpperDiff = "Upper", MiddleDiff = "Middle", LowerDiff = "Lower")) +
  labs(
    title = "RTK Measurement: Change in Sediment Pin Length ",
    subtitle = "From Jun 2022 to Sep 2023   (Positive value = accretion   Negative value = erosion)",
    y = "Change in Pin Length (cm)",
    x = "Marsh Zone",
    fill= "End of Shoreline",
    tag = "RTKpin_boxplots.Rmd_RTKpinchangeEvWandLvNL_box") +
    #caption = "*n=10 west/log, n=7 west/no log, n=12 east/log, n=9 east/no log") +
  theme_bw(base_size = 10) +
  theme(axis.title = element_text(face = "bold", size=13),
        legend.title = element_text(face = "bold", size=09),
        legend.text=element_text(size=09),
        legend.position = "top",   #for no legend type: "none"
        legend.background = element_rect(fill="lightgrey",
                                  size=0.5, linetype="solid", 
                                  colour ="darkgrey"),
        axis.text.x = element_text(size = 11),
        axis.text.y = element_text(size = 10),
        plot.title = element_text(face = "bold", size = 14),
        plot.caption = element_text(size = 09),
        plot.subtitle = element_text(size = 11),
        plot.tag = element_text(size = 6, angle = 90),
        plot.tag.position = "right",
        strip.text = element_text(size = 13, color = "black"))
  
  

RTKpinchangeEvWandLvNL_box
#Will say: Warning: Removed 3 rows containing non-finite outside the scale range (`stat_boxplot()`). This is due to having NA's for a few of the pins.  
```


