---
title: "Thesis-Vegetation Transects"
output: html_notebook
---

RUN! Check that you're using the most updated data sheet and if not, update it in the folder listed below!
```{r}
setwd("/Users/bmorr/OneDrive/Documents/Sears Point Project/R/Thesis")
getwd()
```
Set R up to import excel files with multiple sheets -- install package xlsx
```{r Run}
install.packages("readxl") 
library("readxl")
library("tidyverse")
library("dplyr")
library("ggplot2")
library("hrbrthemes")
```
We have a function in R called read_excel() which we will use to import specific sheet into R. If no argument is specified, then the read_excel() will by default import the first Excel sheet.

RUN!  Cheatsheet: https://github.com/rstudio/cheatsheets/blob/main/data-import.pdf
```{r RUN}
vegdata <- read_excel("VegData.xlsx")
```

RUN! Time to clean up the data:
First, BoMa and other columns read as "logical". Set to numeric:
```{r}
#one at a time
vegdata$BoMa <- as.numeric(vegdata$BoMa)
#multiple at a time
vegdata <- vegdata %>% 
  mutate_at(c('1_MeIn', '2_DiGr', '4_RaSa', '5_CaMa', '6_SaSo', '7_Lolium', '9_FrSa', '15_Fabiaceae','20_Vicia', '21_Sonc', '22_Xant', '23_LeTr', 'Coytote_Bush'), as.numeric)

head(vegdata)
str(vegdata)

vegdata <- vegdata %>% add_column("Date"=paste0(vegdata$Month, "_", vegdata$Year))


```

!Convert decimal values for percent values for the Percent_Spartina:Percent_Bare_ground columns:
```{r DON'T NEED!}
install.packages("scales")                                   
library("scales")

vegdata$Percent_Spartina <- scales::percent(vegdata$Percent_Spartina)
vegdata$Percent_Pickleweed <- scales::percent(vegdata$Percent_Pickleweed)
vegdata$Percent_Upland <- scales::percent(vegdata$Percent_Upland)
vegdata$Percent_Bare_Ground <- scales::percent(vegdata$Percent_Bare_Ground)
```

#From Margot Lessons - Session4_BeccaVegTransect: Remove BaGr Types 1 and 2 and coytote Bush. Since we aren't differentiating bw the two bare ground types we should just combine those two columns into one new column and then remove them:
```{r Do we need to do??!}
vegdata <- vegdata %>%
  select(-BaGr_Type_1, -BaGr_Type_2, -Coytote_Bush) %>% 
  mutate(BaGr = rowSums(across(BaGr_1:BaGr_2), na.rm = T)) %>% 
  select(-BaGr_1, -BaGr_2) 
  
```


!Create a data frame using a small chunk of the data:
```{r DON'T RUN!}
testdata1 <- data.frame(
  Date = rep("February 2023", 11),
  Transect_ID = c("C1-REF1", "C1-REF1", "C1-REF1", "C1-REF1", "C1-REF1", "C1-REF1",
    "C1-REF1", "C1-REF1", "C1-REF1", "C1-REF1", "C1-REF1"),
  Distance_meters = c(0, 1, 2, 3, 4, 5, 6, 8, 10, 12, 14),
  Quadrat = 1:11,
  Zone = c("U", "U", "U", "M", "M", "M", "L", "L", "L", "L", "L"),
  Percent_Spartina = c(
    "0%", "0%", "0%", "0%", "0%", "0%", "0%", "0%", "0%", "0%", "40%"),
  Percent_Pickleweed = c(
    "48%", "88%", "92%", "56%", "80%", "40%", "16%", "100%", "100%", "0%", "32%"),
  Percent_Upland = c(
    "96%", "100%", "12%", "92%", "0%", "0%", "0%", "0%", "0%", "0%", "0%")) %>%
  
#Need to change percent to a numerical value!!! Are these percentages right? They represent a percentage of 25 quadrats. In Margot lessons we multiply by 4 bc 15 is 1/4 of 100.
as.numeric(sub("%","",))/100


# Assuming your data frame is called 'data'


testdata1$Dominant_Type <- apply(testdata1[, c("Percent_Spartina", "Percent_Pickleweed", "Percent_Upland")], 1, function(x) {
  types <- c("Percent_Spartina", "Percent_Pickleweed", "Percent_Upland")
  Dominant_Type <- types[which.max(x)]
  return(Dominant_Type)
})

```
!Giving plotting a try! 
```{r DON'T RUN}
ggplot(testdata1, aes(x = Distance_meters,y = Transect_ID, fill = Percent_Pickleweed)) +
  geom_tile() +
  theme_bw() +
  labs(x = "Distance (meters)", y = "Transect ID", fill = "Percent_Pickleweed") +
  theme(legend.position = "bottom")
```
```{r DON'T RUN}
testdata1 %>%
ggplot(aes(x = Distance_meters,y = Transect_ID,
          group = Quadrat,
          color = Veg_Community)) +
  geom_line() 
```

!Create a dataframe with all of the data:
```{r  DON'T RUN}
testdata2 <- vegdata %>%
  select(Month, Year, Transect_ID, Distance_meters, Percent_Spartina, Percent_Pickleweed, Percent_Upland) %>%
  pivot_longer(cols = Percent_Spartina:Percent_Upland,
                names_to = "Veg_Community",
                values_to = "Percent_cover") %>% 
  group_by(Transect_ID, Month, Year, Distance_meters) %>%
  mutate(Month = factor(Month, levels = c("February", "April", "June", "July", "August", "September", "October")),
           Year = factor(Year, levels = c("2022", "2023"))) 
   

testdata2$Dominant_Type <- apply(testdata2[, c("Percent_Spartina", "Percent_Pickleweed", "Percent_Upland")], 1, function(x) {
  types <- c("Percent_Spartina", "Percent_Pickleweed", "Percent_Upland")
  Dominant_Type <- types[which.max(x)]
  return(Dominant_Type)
})

  
  testdata2 %>%
ggplot(aes(x = Distance_meters,y = Month,
          group = Month,
          color = Veg_Community)) +
  geom_line() 
```




RUN THIS OR THE CHUNK BELOW! BOTH NOT NECESARY.  Same way as chunk below but using for loops:
```{r}
vegdata<-data.frame(vegdata)
str(vegdata)

#confirming that the loop is working on the data you want it to
for (i in 1:nrow(vegdata)) {print(i)
}

for (i in 1:nrow(vegdata)) {max(vegdata$Percent_Spartina[i], vegdata$Percent_Pickleweed[i], vegdata$Percent_Upland[i])
}

vegdata$Dominant_Community <- max(vegdata$Percent_Spartina[i], vegdata$Percent_Pickleweed[i], vegdata$Percent_Upland[i], vegdata$Percent_Bare_Ground[i])

vegdata$Dominant_Community <- NA
vegdata$Dominant_Community <- as.numeric("NULL")

for (i in 1:nrow(vegdata)) {
  vegdata$Dominant_Community [i] <- max(vegdata$Percent_Spartina[i], vegdata$Percent_Pickleweed[i], vegdata$Percent_Upland[i], vegdata$Percent_Bare_Ground[i])
}

```

Same way as above but using Tidyverse:
```{r}
vegdata %>% 
  mutate(Dom_Community_ID = case_when(Dominant_Community == Percent_Spartina ~ "Spartina", 
                                      Dominant_Community == Percent_Pickleweed ~ "Pickleweed",
                                      Dominant_Community == Percent_Upland ~ "Upland",
                                      Dominant_Community == Percent_Bare_Ground ~ "Bare_ground"))


select(Transect_ID,Date, Quadrat, Percent_Spartina:Percent_Upland, Percent_Bare_Ground) %>% 
  pivot_longer(cols = -c("Transect_ID", "Date", "Quadrat"), names_to = "Dominant_Community")  %>% 
  mutate(Dominant_Community = str_remove(Dominant_Community, "Percent_")) %>% 
  group_by(Transect_ID, Date, Quadrat) %>% 
  summarise(Percent_Cover = max(value, na.rm=TRUE), Dominant_Community = Dominant_Community) %>% 
            
  
```



Now I want to make a new data frame that only has the columns I need for these plots:
```{r}
stuartplots <- vegdata %>% 
   select(-SpFo:-X23_LeTr)
```

Start working with the data for lollipop charts from here: https://r-graph-gallery.com/303-lollipop-plot-with-2-values.html
```{r}
#here's the provided code:

ggplot(data) +
  geom_segment( aes(x=x, xend=x, y=value1, yend=value2), color="grey") +
  geom_point( aes(x=x, y=value1), color=rgb(0.2,0.7,0.1,0.5), size=3 ) +
  geom_point( aes(x=x, y=value2), color=rgb(0.7,0.2,0.1,0.5), size=3 ) +
  coord_flip()+
  theme_ipsum() +
  theme(
    legend.position = "none",
  ) +
  xlab("") +
  ylab("Value of Y")

#my attempt:

ggplot(stuartplots) +
  geom_segment(aes(x= Distance_meters, xend=Distance_meters, y=Transect_ID, yend=Transect_ID), color="black") +
  geom_segment(aes(x= Distance_meters, xend=Distance_meters, y=Transect_ID, yend=Transect_ID), color="black") +
  xlab("Distance Along Transect") +
  ylab("Transect_ID")

```








