---
title: "VegData_stats"
author: "Becca Morris"
date: "2024-02-03"
output: html_document
---

```{r setup, include=FALSE}
library(readxl)
library(tidyverse)
library(ggplot2)
library(broom)
library(AICcmodavg)
library(glmmTMB)
library(lme4)
library(bbmle)
library(mvabund)  #allows you to run meanvar.plot() 


setwd("~/GitHub/BeccaTransects/data")

vegmetadata <- read_xlsx("~/GitHub/BeccaTransects/data/VegData_raw.xlsx", sheet = "Metadata")
vegdata <- read_xlsx("~/GitHub/BeccaTransects/data/VegData_raw.xlsx", sheet = "Data_2")
```

#!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!#
# 1. creating the main data frame:
```{r}
vegdata2 <- vegdata %>% 
  mutate_at(c('MeIn', 'DiGr', 'RaSa', 'CaMa', 'SaSo', 'Lolium', 'FrSa', 'Fabiaceae','Vicia', 'Sonc', 'Xant', 'LeTr', 'Ranunculus', 'Taraxacum', 'Coytote_Bush','BoMa', 'DiSp', 'AtPx', 'PoAr', 'AcMa', 'GrSt', 'MixedSeedlings','GrGR', 'Melilotus', 'MePo'), as.numeric) %>% 
  mutate_if(is.numeric, list(~replace_na(., 0))) %>%
  filter(!(BaGr_Type_1 == '7' | BaGr_Type_2 == '7')) %>% #remove rows where BaGr is coded as log
  select(-BaGr_Type_1, -BaGr_Type_2, -Coytote_Bush, -Wrack, -Tot_Num_Species, -Shoot_Density, -Length_centimeters, -Patch_Type) %>% 
  mutate(SaSp = SaPa + SaDe) %>% 
  mutate(SaPa_new = SaPa + ((SaPa/SaSp)*PicklePups)) %>% 
  mutate(SaDe_new = SaDe + ((SaDe/SaSp)*PicklePups)) %>% 
  mutate(Grass = GrGR + LeTr + MixedSeedlings) %>% 
  mutate(BaGr = rowSums(across(BaGr_1:BaGr_2), na.rm = T)) %>% 
  select(-BaGr_1, -BaGr_2) %>% 
  mutate(Total_Veg = 25-BaGr) %>%  #Total_Veg is simply what is remaining after subtracting the bare ground
  #mutate(across(SpFo:Total_Veg, ~ .x*4)) %>%  #makes percentages out of the counts
  mutate(across(everything(), ~replace(.x, is.nan(.x), 0))) %>%
  select(-c('GrGR','LeTr','MixedSeedlings','PicklePups', 'SaPa','SaDe', 'SaSp')) %>%  #removed SaSp so it would not be calculated as total veg in addition tot he SaPa and SaDe columns)) %>% 
  rename(SaPa = SaPa_new, SaDe = SaDe_new) %>% #putting their original names back after the calculation above
  mutate(Total_Veg2 = rowSums(across(SpFo:Grass), na.rm =T)) %>% #Total_Veg2 adds all the total quadrants each species is present in and combining them all into one total count (can exceed 100)
  mutate(Date = paste0(Year, "-", Month)) 
  
  
  #filter(!Total_Veg2 == 0)  #remove all rows where total_veg is equal to 0
  
```

#### Using Full_Join to combine main data sheet and metadata sheet ####

# 2. First I want to clean up the metadata sheet:
```{r}
vegmetadata %>% 
  select(Transect_ID, Shoreline_End, Log_Presence)
```

# 3. Now to merge the two data frames together, combine pickleweed species into a single column, distirbute picklepups into SaPa and SaDe columns, combine grasses into one column, and remove the redunant columns:
```{r}
vegdatamerged <- full_join(vegdata2, vegmetadata, by = "Transect_ID") %>% 
  select(-c(...9:...12)) %>% 
  mutate(Shoreline_End = factor(Shoreline_End, levels = c("west", "east")), #need to set these as factor to run a glmm
        Zone = factor(Zone, levels = c("U", "M", "L")),
         Log_Presence = factor(Log_Presence, levels = c("log", "no log")))
```


#GLMM attempt:

#since the same transects are measured repeatedly, the transects are used as a random effect
```{r}
hist(vegdatamerged$Total_Veg2)
#QQ-Plot:
qqnorm(vegdatamerged$Total_Veg2, pch = 1, frame = FALSE)
qqline(vegdatamerged$Total_Veg2, col = "steelblue", lwd = 2)

#attempting to use mvabund to make a mean vs variance plot ##cant figure out how to work yet
tikus <- data("tikus")
vegdata_mvabund <- mvabund(vegdatamerged$Total_Veg2)
meanvar.plot(vegdata_mvabund)
```



# 4. Now to break the data up into the three zones and remove the reference logs. 
# *The middle zone contains logs, so I'll need to make the log location's NA's:

# Upper Zone:
```{r}
vegdata_U <- vegdatamerged[(vegdatamerged$Zone == "U" & vegdatamerged$Log_Presence != "reference"),]  %>% 
  filter(Date == "2023-August") 


#calculating change in cover for total veg (where total can exceed 25) and then the mean change for each zone, shoreline end, and log presence:
vegdata_U <- vegdatamerged[(vegdatamerged$Zone == "U" & vegdatamerged$Log_Presence != "reference"),] %>% 
  filter(Date == "2022-August" | Date == "2023-August") %>% 
  group_by(Transect_ID, Quadrat) %>% 
  arrange(Date) %>% #arrange by date so the calculations can be made below
  mutate(initial_count = first(Total_Veg2),
         final_count = last(Total_Veg2),
         count_difference = (final_count - initial_count)) %>%
  ungroup() %>% 
  filter(!(Date == "2022-August")) %>%  #this just removes the June count_difference data bc it is a repeat for the august data and we only need one calculation
  group_by(Date, Zone, Shoreline_End, Log_Presence) %>% 
  summarize(mean_cover = mean(count_difference, na.rm = TRUE),
            sd_cover = sd(count_difference, na.rm = TRUE),
            n_cover = n()) %>% 
  mutate(se_cover = sd_cover/sqrt(n_cover)) %>% 
  mutate(across(where(is.numeric), ~round(., 0))) %>% 
  mutate(Shoreline_End = factor(Shoreline_End, levels = c("west", "east")),
        Zone = factor(Zone, levels = c("U", "M", "L")),
         Log_Presence = factor(Log_Presence, levels = c("log", "no log")))
```
# Middle Zone:
```{r}
vegdata_M <- vegdatamerged[(vegdatamerged$Zone == "M" & vegdatamerged$Log_Presence != "reference"),] %>% 
  filter(Date == "2023-August") 

vegdata_M <- vegdatamerged[(vegdatamerged$Zone == "M" & vegdatamerged$Log_Presence != "reference"),] %>% 
  filter(Date == "2022-August" | Date == "2023-August") %>% 
  group_by(Transect_ID, Quadrat) %>% 
  arrange(Date) %>% #arrange by date so the calculations can be made below
  mutate(initial_count = first(Total_Veg2),
         final_count = last(Total_Veg2),
         count_difference = (final_count - initial_count)) %>%
  ungroup() %>% 
  filter(!(Date == "2022-August")) %>%  #this just removes the June count_difference data bc it is a repeat for the august data and we only need one calculation
  group_by(Date, Zone, Shoreline_End, Log_Presence) %>% 
  summarize(mean_cover = mean(count_difference, na.rm = TRUE),   # !!!!!!!!!!!!!   from here down you calculate summary statistics
            sd_cover = sd(count_difference, na.rm = TRUE),
            n_cover = n()) %>% 
  mutate(se_cover = sd_cover/sqrt(n_cover)) %>% 
  mutate(across(where(is.numeric), ~round(., 0))) %>% 
  mutate(Shoreline_End = factor(Shoreline_End, levels = c("west", "east")),
        Zone = factor(Zone, levels = c("U", "M", "L")),
         Log_Presence = factor(Log_Presence, levels = c("log", "no log")))
```
# Lower Zone:
```{r}
vegdata_L <- vegdatamerged[(vegdatamerged$Zone == "L" & vegdatamerged$Log_Presence != "reference"),] %>% 
  filter(Date == "2023-August") 

vegdata_L <- vegdatamerged[(vegdatamerged$Zone == "L" & vegdatamerged$Log_Presence != "reference"),] %>% 
  filter(Date == "2022-August" | Date == "2023-August") %>% 
  group_by(Transect_ID, Quadrat) %>% 
  arrange(Date) %>% #arrange by date so the calculations can be made below
  mutate(initial_count = first(Total_Veg2),
         final_count = last(Total_Veg2),
         count_difference = (final_count - initial_count)) %>%
  ungroup() %>% 
  filter(!(Date == "2022-August")) %>%  #this just removes the June count_difference data bc it is a repeat for the august data and we only need one calculation
  group_by(Date, Zone, Shoreline_End, Log_Presence) %>% 
  summarize(mean_cover = mean(count_difference, na.rm = TRUE),   #from here down you calculate summary statistics
            sd_cover = sd(count_difference, na.rm = TRUE),
            n_cover = n()) %>% 
  mutate(se_cover = sd_cover/sqrt(n_cover)) %>% 
  mutate(across(where(is.numeric), ~round(., 0))) %>% 
  mutate(Shoreline_End = factor(Shoreline_End, levels = c("west", "east")),
        Zone = factor(Zone, levels = c("U", "M", "L")),
         Log_Presence = factor(Log_Presence, levels = c("log", "no log")))
```



#!!!!!!!!!!!!!!!!!!!!!!#
# 5. Test for normality:

# Upper Zone Normality using the Total_Veg2 column created in step 1:
```{r}
#change variable based on which dataset youare using!!

#QQ-Plot:
qqnorm(vegdata_U$count_difference, pch = 1, frame = FALSE)
qqline(vegdata_U$count_difference, col = "steelblue", lwd = 2)

#histogram
hist(vegdata_U$Total_Veg2) 

#since SaPa is a combo of Pickle Pups and the original SaPa column, I need to tell it to not exceed 25 where Sapa is > than 25.
vegdata_U$SaPa[vegdata_U$SaPa > 25] <- 25
hist(vegdata_U$SaPa)


```
# Middle Zone Normality:
```{r}
#QQ-Plot:
qqnorm(vegdata_M$count_difference, pch = 1, frame = FALSE)
qqline(vegdata_M$count_difference, col = "steelblue", lwd = 2)

#histogram
hist(vegdata_M$Total_Veg2)

```
# Lower Zone Normality:
```{r}
#QQ-Plot:
qqnorm(vegdata_L$count_difference, pch = 1, frame = FALSE)
qqline(vegdata_L$count_difference, col = "steelblue", lwd = 2)

#histogram
hist(vegdata_L$Total_Veg2)
```

#### ANOVA: tests whether any of the group means are different from the overall mean of the data by checking the variance of each individual group against the overall variance of the data. If one or more groups falls outside the range of variation predicted by the null hypothesis (all group means are equal), then the test is statistically significant. 
```{r}

```

#### General GLM:  NO GOOD FOR MY DATA

#Upper Zone GLM:
```{r}
hist((vegdata_U$Total_Veg2[vegdata_U$Date == "2023-August"]))

glm1 <- glm((Total_Veg2) ~ Shoreline_End*Log_Presence,
               data = vegdata_U[vegdata_U$Date == "2023-August",],
               family = gaussian)

glm1
plot(glm1)
anova(glm1, test = "F")

```






# ################################################################################################################################ #


#Binomial GLM

#Analyzing bare ground vs Total Veg using the "Total_Veg" and BaGr" columns and a binomial glm with logit/log odds description. Good for non-normal data:

#Question: Does Shoreline_End*Log_Presence affect how much veg/bare ground is present? 

#Vegetation:
```{r}
#need to replace zero's with NA:
vegdata_U <- vegdatamerged[(vegdatamerged$Zone == "U" & vegdatamerged$Log_Presence != "reference"),] %>% 
   mutate(Vegetation = +(Total_Veg>=1))
  
  
vegdata_M <- vegdatamerged[(vegdatamerged$Zone == "M" & vegdatamerged$Log_Presence != "reference"),] %>% 
   mutate(Vegetation = +(Total_Veg>=1))
  

vegdata_L <- vegdatamerged[(vegdatamerged$Zone == "L" & vegdatamerged$Log_Presence != "reference"),] %>% 
   mutate(Vegetation = +(Total_Veg>=1))
  
```
#Upper Zone:


#Running the glm with a binomial distribution:
#First, a complex model taking interactions into account:
```{r}
BiVegGLM_U1 <- glm(Vegetation ~ Shoreline_End*Log_Presence, family=binomial (link = "logit"), data=vegdata_U)
plot(BiVegGLM_U1)
```
#Next, a simpler model for only main effects:
```{r}
BiVegGLM_U2 <- glm(Vegetation ~ Shoreline_End+Log_Presence, family=binomial (link = "logit"), data=vegdata_U)
plot(BiVegGLM_U2)
```
#Now, comparing the two models using a chi^2 test (Contingency test):
```{r}
#https://www.youtube.com/watch?v=KubK6hgMbvg

#Chi^2 hypothesis:
#Ho: No relationship exists on the categorical variables in the population; categorical variables are independent.Knowing the value of one DOES NOT predict the other.
#Hi: Categorical variables are not independent. Knowing the value of one DOES predict the other variable.

anova(BiVegGLM_U1, BiVegGLM_U2, test = "Chi") 
#Chi^2 Results: 
#Model "BiVegGLM_U2" has a p-value = 0.08159, so we model 2 is more appropriate. 

# ***The categorical variables (shoreline end and log presence) ARE independent of each other. ie. Use model 2.
```

```{r}
summary(BiVegGLM_U2)$coefficients
#Shoreline end and log presence main effects are both significant
#Explanation of results: https://www.simonqueenborough.info/R/statistics/glm-binomial:
#For categorical predictors, the coefficient estimate describes the change in the log odds for each level compared to the base level.

# Add logistic fitted values back to dataframe as
#  new column pred.vegetation
```
#Now creating confidence intervals in case I need them:
```{r}
confint(BiVegGLM_U2)
```



#Middle Binomial GLM:

#Running the glm with a binomial distribution:
#First, a complex model taking interactions into account:
```{r}
BiVegGLM_M1 <- glm(Vegetation ~ Shoreline_End*Log_Presence, family=binomial (link = "logit"), data=vegdata_M)
```
#Next, a simpler model for only main effects:
```{r}
BiVegGLM_M2 <- glm(Vegetation ~ Shoreline_End+Log_Presence, family=binomial (link = "logit"), data=vegdata_M)
```
#Now, comparing the two models using a chi^2 test (Contingency test):
```{r}
#Chi^2 hypothesis:
#Ho: No relationship exists on the categorical variables in the population; categorical variables are independent. 
#Hi: Categorical variables are not independent. 

anova(BiVegGLM_M1, BiVegGLM_M2, test = "Chi") 
#Chi^2 Results: 
#Model "BiVegGLM_U2" has a p-value= 4.811e-14 ***, which means we accept the null hypothesis.

#*** The categorical variables (shoreline end and log presence) are NOT independent of each other. We must investigate the mixed effects of the two. ie. Use model 2.
```
#Now creating confidence intervals in case I need them:
```{r}
confint(BiVegGLM_M2)
```



# Lower Binomial GLM:

#Running the glm with a binomial distribution:
#First, a complex model taking interactions into account:
```{r}
BiVegGLM_L1 <- glm(Vegetation ~ Shoreline_End*Log_Presence, family=binomial (link = "logit"), data=vegdata_L)
```
#Next, a simpler model for only main effects:
```{r}
BiVegGLM_L2 <- glm(Vegetation ~ Shoreline_End+Log_Presence, family=binomial (link = "logit"), data=vegdata_L)
```
#Now, comparing the two models using a chi^2 test (Contingency test):
```{r}
#Chi^2 hypothesis:
#Ho: No relationship exists on the categorical variables in the population; categorical variables are independent. 
#Hi: Categorical variables are not independent. 

anova(BiVegGLM_L1, BiVegGLM_L2, test = "Chi") 
#Chi^2 Results: 
#Model "BiVegGLM_L2" has a p-value= 0.5867, which means we reject the null hypothesis.

#***The categorical variables (shoreline end and log presence) ARE independent of each other.  ie. Use model 1.
```
#Now creating confidence intervals in case I need them:
```{r}
confint(BiVegGLM_M2)
```



# ############################################################################################################################################################################# #

#Now I want to determine if I have differences in communities within the vegetation that is present:

#need to make new dataframe where 0's are NA's, and add a new column for relative cover:
```{r}
relative <- vegdatamerged %>% 
  mutate(across(c(7:38), ~na_if(., 0))) %>% 
  mutate(rel_SpFo = SpFo/Total_Veg) %>% 
  mutate(rel_SaPa = SaPa/Total_Veg) %>%
  mutate(rel_SaSp = SaSp/Total_Veg)
  mutate(across(where(is.numeric), ~round(., 0))) %>% 
  mutate(across(c(7:38), ~na_if(., 0)))  
  

```


# Upper Zone:
```{r}
vegdata_U <- vegdatamerged[(vegdatamerged$Zone == "U" & vegdatamerged$Log_Presence != "reference"),] %>% 
  
```
# Middle Zone:
```{r}
vegdata_M <- vegdatamerged[(vegdatamerged$Zone == "M" & vegdatamerged$Log_Presence != "reference"),]
```
# Lower Zone:
```{r}
vegdata_L <- vegdatamerged[(vegdatamerged$Zone == "L" & vegdatamerged$Log_Presence != "reference"),]
```
