---
title: "Veg nMDS"
author: "Becca Morris"
date: "2024-02-06"
output: html_document
---

```{r}
library(readxl)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(vegan)



setwd("~/GitHub/BeccaTransects/data")

vegmetadata <- read_xlsx("~/GitHub/BeccaTransects/data/VegData_raw.xlsx", sheet = "Metadata")
vegdata <- read_xlsx("~/GitHub/BeccaTransects/data/VegData_raw.xlsx", sheet = "Data")
```

# 1. creating the main data frame:
```{r}

VegDataNew <- vegdata %>% 
  mutate_at(c('MeIn', 'DiGr', 'RaSa', 'CaMa', 'SaSo', 'Lolium', 'FrSa', 'Fabiaceae','Vicia', 'Sonc', 'Xant', 'LeTr', 'Ranunculus','Taraxacum', 'Coytote_Bush','BoMa'), as.numeric) %>% 
  mutate_if(is.numeric, list(~replace_na(., 0))) %>%
  select(-BaGr_Type_1, -BaGr_Type_2, -Coytote_Bush, -Wrack, -Tot_Num_Species, -Shoot_Density, -Length_centimeters,    -Patch_Type) %>% 
  mutate(BaGr = rowSums(across(BaGr_1:BaGr_2), na.rm = T)) %>% 
  select(-BaGr_1, -BaGr_2) %>% 
  mutate(Total_Veg = 25-BaGr) %>%  #Total_veg is simply what is remaining after subtracting the bare ground
  mutate(across(SpFo:Total_Veg, ~ .x*4)) %>% 
  mutate(Date = paste0(Year, "-", Month)) %>% 
  mutate(Total_Veg2 = rowSums(across(SpFo:LeTr))) %>% 
  filter(!Total_Veg2 == 0) 

VegDataMerged <- full_join(VegDataNew, vegmetadata, by = "Transect_ID") %>% 
  select(-c(...9:...12)) %>% 
  mutate(SaSp = SaPa + SaDe) %>% 
  mutate(SaPa_new = SaPa + ((SaPa/SaSp)*PicklePups)) %>% 
  mutate(SaDe_new = SaDe + ((SaDe/SaSp)*PicklePups)) %>% 
  mutate(Grass = GrGR + LeTr + MixedSeedlings) %>% 
  mutate(across(everything(), ~replace(.x, is.nan(.x), 0)))
 



```

#2. Now I need to seperate my data into two df's for EACH ZONE -- one for identifiers and one for abundance data only.
# *We removed the row where C6-L174 is equal to quadrat 3 and zone U since it was a big outlier.
```{r}
VegDataID_U <- VegDataMerged[VegDataMerged$Date == "2023-August" & VegDataMerged$Zone == "U" &
                             !(VegDataMerged$Transect_ID == "C6-L174" & VegDataMerged$Quadrat == 3 &   VegDataMerged$Zone == "U"),
                           c('Transect_ID', 'Distance_meters', 'Quadrat', 'Zone', 'Date', 'Shoreline_End', 'Log_Presence')]



VegDataSpecies_U <- VegDataMerged[(VegDataMerged$Date == "2023-August" & VegDataMerged$Zone == "U" &
                                  !(VegDataMerged$Transect_ID == "C6-L174" & 
                                      VegDataMerged$Quadrat == 3)), 
                                c(7:37, 50:52)] %>% 
  select(-c('GrGR','LeTr','MixedSeedlings','SaPa','SaDe','PicklePups')) #'CaMa', 'Lolium', 'RaSa', 'FrSa', 'Fabiaceae', 'Ranunculus', 'Taraxacum'))



#Transform the veg communities using the hellinger transformation.This transformation is useful for abundance data when you may have a few really abundant species and many rare species.It transforms the abundances of the really abundant and few rare species don't have less of an influence on the ordination. hellinger: square root of method = "total" (Legendre & Gallagher 2001).

VegDataSpecies.hel_U <- decostand(VegDataSpecies_U, method = "hellinger")


```

```{r}
VegDataID_M <- VegDataMerged[VegDataMerged$Date == "2023-August" & VegDataMerged$Zone == "M",
                             #!(VegDataMerged$Transect_ID == "C6-L174" & VegDataMerged$Quadrat == 3 & VegDataMerged$Zone == "U"),
                           c('Transect_ID', 'Distance_meters', 'Quadrat', 'Zone', 'Date', 'Shoreline_End', 'Log_Presence')]


VegDataSpecies_M <- VegDataMerged[VegDataMerged$Date == "2023-August" & VegDataMerged$Zone == "M",
                                  #!(VegDataMerged$Transect_ID == "C6-L174" & 
                                      #VegDataMerged$Quadrat == 3)),
                                c(7:37)] %>% 
  select(-c('CaMa', 'Lolium', 'RaSa', 'FrSa', 'Fabiaceae', 'Ranunculus', 'Taraxacum'))
```

```{r}
VegDataID_L <- VegDataMerged[VegDataMerged$Date == "2023-August" & VegDataMerged$Zone == "L",
                             #!(VegDataMerged$Transect_ID == "C6-L174" & VegDataMerged$Quadrat == 3 & VegDataMerged$Zone == "U"),
                           c('Transect_ID', 'Distance_meters', 'Quadrat', 'Zone', 'Date', 'Shoreline_End', 'Log_Presence')]


VegDataSpecies_L <- VegDataMerged[VegDataMerged$Date == "2023-August" & VegDataMerged$Zone == "L",
                                  #!(VegDataMerged$Transect_ID == "C6-L174" & 
                                      #VegDataMerged$Quadrat == 3)),
                                c(7:37)] %>% 
  select(-c('CaMa', 'Lolium', 'RaSa', 'FrSa', 'Fabiaceae', 'Ranunculus', 'Taraxacum'))
```

# Per Margot: Run the nMDS. This is where I previously had an extra step, which I removed. You can play with the value of k, which is the number of dimensions. #2 is best but if you increase dimensions to 3 you decrease stress, which might be necessary for high values.
#Stress: 0-0.1 = excellent fit, 0.1 – 0.2 = good fit, 0.2 – 0.3 = very mediocre fit, 0.3+ = trash
#The default is trymax = 20. This is often not enough to achieve convergence, so I often increase this value
#I generally don’t increase it past 10,000 – anything more than that and your data is not going to work with the analysis, or you need to mess with k
# *** If they are all hashtagged out it is to preserve the last run of the MDS. Everything is all good.
```{r}
### UPPER NMDS ###
set.seed(19840217)
#VegDataSpeciesNMDS_U <- metaMDS(VegDataSpecies_U, k = 2, trymax = 20)
VegDataSpeciesNMDS_U 

#VegDataSpeciesNMDS_U  <- metaMDS(VegDataSpecies_U, previous.best = VegDataSpeciesNMDS_U)




### Hellinger Method ###

nmds_U <- metaMDS(VegDataSpecies.hel_U, autotransform = F, k = 2, trymax = 20) #autotransform = F so that it doesn't revert back to the default transformation *bc we are using the hellinger method.
nmds_U

save(nmds_U, file ="../output_data/nmds_U.Rds")

load("output_data/nmds_U.RData")
```

```{r}
### MIDDLE NMDS ###
set.seed(19840217)
#VegDataSpeciesNMDS_M <- metaMDS(VegDataSpecies_M, k = 2, trymax = 20)
#VegDataSpeciesNMDS_M 

#VegDataSpeciesNMDS_M  <- metaMDS(VegDataSpecies_M, previous.best = VegDataSpeciesNMDS_U)
```

```{r}
### LOWER NMDS ###
set.seed(19840217)
#VegDataSpeciesNMDS_L <- metaMDS(VegDataSpecies_L, k = 2, trymax = 20)
#VegDataSpeciesNMDS_L 

#VegDataSpeciesNMDS_L  <- metaMDS(VegDataSpecies_L, previous.best = VegDataSpeciesNMDS_U)
```


### Plotting ###

#Starting with Upper Zone:
```{r}
#Base R plotting:
stressplot(VegDataSpeciesNMDS_U)

plot(VegDataSpeciesNMDS_U)


#ggplot:
Axis1 <- as.numeric(as.character(VegDataSpeciesNMDS_U$points[,1]))

Axis2 <- as.numeric(as.character(VegDataSpeciesNMDS_U$points[,2]))

mds.coords1_U <- data.frame(VegDataID_U, Axis1, Axis2)

ggplot() +
  geom_point(data = mds.coords1_U, aes(x=Axis1, y=Axis2, 
                                       color = Log_Presence,
                                       shape = Shoreline_End)) +
               labs(title = "Upper Zone nMDS") +
               theme_bw()


###  Plotting Hellinger method in base R ###

ordiplot(nmds_U)
ordiplot(nmds_U, type = "t") #puts species name on plot

# using ggplot with Hellinger Method ###

Axis1 <- as.numeric(as.character(nmds_U$points[,1]))

Axis2 <- as.numeric(as.character(nmds_U$points[,2]))

mds.coords1_U <- data.frame(VegDataID_U, Axis1, Axis2)


ggplot() +
  geom_point(data = mds.coords1_U, aes(x=Axis1, y=Axis2, 
                                       color = Log_Presence,
                                       shape = Shoreline_End)) +
               labs(title = "Upper Zone nMDS") +
               theme_bw() 


```
#Middle Zone
```{r}
#Base R plotting:
stressplot(VegDataSpeciesNMDS_M)

plot(VegDataSpeciesNMDS_M)


#ggplot:
Axis1 <- as.numeric(as.character(VegDataSpeciesNMDS_M$points[,1]))

Axis2 <- as.numeric(as.character(VegDataSpeciesNMDS_M$points[,2]))

mds.coords1_M <- data.frame(VegDataID_M, Axis1, Axis2)

ggplot() +
  geom_point(data = mds.coords1_M, aes(x=Axis1, y=Axis2, 
                                       color = Log_Presence,
                                       shape = Shoreline_End)) +
               labs(title = "Middle Zone nMDS") +
               theme_bw()
```
#Lower Zone:
```{r}
#Base R plotting:
stressplot(VegDataSpeciesNMDS_L)

plot(VegDataSpeciesNMDS_L)


#Margot way:
Axis1 <- as.numeric(as.character(VegDataSpeciesNMDS_L$points[,1]))

Axis2 <- as.numeric(as.character(VegDataSpeciesNMDS_L$points[,2]))

mds.coords1_L <- data.frame(VegDataID_L, Axis1, Axis2)

ggplot() +
  geom_point(data = mds.coords1_L, aes(x=Axis1, y=Axis2, 
                                       color = Log_Presence,
                                       shape = Shoreline_End)) +
               labs(title = "Lower Zone nMDS") +
               theme_bw()
```

