---
title: "Veg_stats_ZANB"
author: "Becca Morris"
date: "2024-02-27"
output: html_document
---
```{r setup, include=FALSE}
library(readxl)
library(tidyverse)
library(ggplot2)
library(broom)
library(AICcmodavg)
library(glmmTMB)
library(lme4)
library(bbmle)
library(pscl)
library(lmtest) #runs a likelyhood ratio test


setwd("~/GitHub/BeccaTransects/data")

vegmetadata <- read_xlsx("~/GitHub/BeccaTransects/data/VegData_raw.xlsx", sheet = "Metadata")
vegdata <- read_xlsx("~/GitHub/BeccaTransects/data/VegData_raw.xlsx", sheet = "Data_2")
```

#!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!#
# 1. creating the main data frame:
```{r}
vegdata2 <- vegdata %>% 
  mutate_at(c('MeIn', 'DiGr', 'RaSa', 'CaMa', 'SaSo', 'Lolium', 'FrSa', 'Fabiaceae','Vicia', 'Sonc', 'Xant', 'LeTr', 'Ranunculus', 'Taraxacum', 'Coytote_Bush','BoMa', 'DiSp', 'AtPx', 'PoAr', 'AcMa', 'GrSt', 'MixedSeedlings','GrGR', 'Melilotus', 'MePo'), as.numeric) %>% 
  mutate_if(is.numeric, list(~replace_na(., 0))) %>%
  filter(!(BaGr_Type_1 == '7' | BaGr_Type_2 == '7')) %>% #remove rows where BaGr is coded as log
  select(-BaGr_Type_1, -BaGr_Type_2, -Coytote_Bush, -Wrack, -Tot_Num_Species, -Shoot_Density, -Length_centimeters, -Patch_Type) %>% 
  mutate(SaSp = SaPa + SaDe) %>% 
  mutate(SaPa_new = SaPa + ((SaPa/SaSp)*PicklePups)) %>% 
  mutate(SaDe_new = SaDe + ((SaDe/SaSp)*PicklePups)) %>% 
  mutate(Grass = GrGR + LeTr + MixedSeedlings) %>% 
  mutate(BaGr = rowSums(across(BaGr_1:BaGr_2), na.rm = T)) %>% 
  select(-BaGr_1, -BaGr_2) %>% 
  mutate(Total_Veg = 25-BaGr) %>%  #Total_Veg is simply what is remaining after subtracting the bare ground
  #mutate(across(SpFo:Total_Veg, ~ .x*4)) %>%  #makes percentages out of the counts
  mutate(across(everything(), ~replace(.x, is.nan(.x), 0))) %>%
  select(-c('GrGR','LeTr','MixedSeedlings','PicklePups', 'SaPa','SaDe', 'SaSp')) %>%  #removed SaSp so it would not be calculated as total veg in addition tot he SaPa and SaDe columns)) %>% 
  rename(SaPa = SaPa_new, SaDe = SaDe_new) %>% #putting their original names back after the calculation above
  mutate(Grass = ifelse(Grass > 25, 25, Grass)) %>% 
  mutate(Total_Veg2 = rowSums(across(SpFo:Grass), na.rm =T)) %>% #Total_Veg2 adds all the total quadrants each species is present in and combining them all into one total count (can exceed 100)
  mutate(Date = paste0(Year, "-", Month)) 
  #mutate(across(c(7:37), ~na_if(., 0))) #Just hashtag out if prefer 0's in place fo NA
  #filter(!Total_Veg2 == 0)  #remove all rows where total_veg is equal to 0
  
```

#### Using Full_Join to combine main data sheet and metadata sheet ####

# 2. First I want to clean up the metadata sheet:
```{r}
vegmetadata %>% 
  select(Transect_ID, Shoreline_End, Log_Presence)
```

# 3. Now to merge the two data frames together, combine pickleweed species into a single column, distirbute picklepups into SaPa and SaDe columns, combine grasses into one column, and remove the redunant columns:
```{r}
vegdatamerged <- full_join(vegdata2, vegmetadata, by = "Transect_ID") %>% 
  select(-c(...9:...12)) %>% 
  mutate(Shoreline_End = factor(Shoreline_End, levels = c("west", "east")), #need to set these as factor to run a glmm
        Zone = factor(Zone, levels = c("U", "M", "L")),
         Log_Presence = factor(Log_Presence, levels = c("log", "no log")))
```


#GLMM attempt:

#since the same transects are measured repeatedly, the transects are used as a random effect
```{r}
hist(vegdatamerged$Total_Veg2) #data is majorly zero-inflated and overdispersed at zero adn around 25
#QQ-Plot:
qqnorm(vegdatamerged$Total_Veg2, pch = 1, frame = FALSE)
qqline(vegdatamerged$Total_Veg2, col = "steelblue", lwd = 2)
```


#ZANB:

#set up a df for this:

#all of this is found starting on pg. 278 in the Mixed Effect model book

# If your explanatory avriables aren't already, set them as factors. *Already done by code up above.
#Set 0's to NA if nto already done. *Already done by code up above
```{r}
I1 <- is.na(vegdatamerged$Total_Veg2) |
  is.na(vegdatamerged$Shoreline_End) |
  is.na(vegdatamerged$Zone) |
  is.na(vegdatamerged$Log_Presence)

vegdatamerged2 <- vegdatamerged[!I1,]
```

```{r}
plot(table(vegdatamerged2$Total_Veg2),
     xlab = "Observed Total Veg Count Values",
     ylab = "Frequencies")
```
```{r}
library(pscl)
```

#Compare the ZAP and ZANB model to detrmine which one is a better:
```{r}
#Applies a zero-inflated model
f1 <- formula(Total_Veg2 ~ Shoreline_End+Zone+Log_Presence + Shoreline_End*Zone*Log_Presence | Shoreline_End+Zone+Log_Presence + Shoreline_End*Zone*Log_Presence) ###### IS this the correct formula?!?!

#Determine which model you need:   *we are testing whether the variance structure of the Poisson is the same as the var structure of the Nb model
Nb1 <- zeroinfl(f1,dist = "negbin", link = "logit",  #ZANB
                data = vegdatamerged2)
Zap1 <- zeroinfl(f1,dist = "poisson", link = "logit",  #ZAP
                data = vegdatamerged2)

lrtest(Nb1, Zap1) #Chi^2 = 2.2 x e-16 *** so use the ZANB model?
summary(Nb1)
```

```{r}
#Another way to determine if you need a ZAP or ZANB model:
H1A <- hurdle(f1, dist = "poisson", link = "logit",  #ZAP
              data = vegdatamerged2)
H1B <- hurdle(f1, dist = "negbin", link = "logit",  #ZANB
              data = vegdatamerged2)

lrtest(H1A, H1B)   #Ho: use the Zap model. Hi: Use the ZANB model *Likelihood ratio test provides Chi^2 value of 2.2 X e-16 *** so I should use the ZANB model
AIC(H1A, H1B)     #AIC H1A = 72059.25 and H1B = 54331.17    *Choose model with the lower AIC value
```

