---
title: "Veg nMDS June 2022"
author: "Becca Morris"
date: "2024-02-21"
output: html_document
---

```{r}
library(readxl)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(vegan)
#everything below is to install the latest version of ggvegan for plotting the nmds
#if (!require(devtools)) {           
#  install.packages("devtools")
#}
#devtools::install_github("gavinsimpson/ggvegan")

library(ggvegan)

#everything below is to install the latest version of ggpubr for plotting the nmds
#if(!require(devtools)) install.packages("devtools")
#devtools::install_github("kassambara/ggpubr")

library(ggpubr)

setwd("~/GitHub/BeccaTransects")

vegmetadata <- read_xlsx("~/GitHub/BeccaTransects/data/VegData_raw.xlsx", sheet = "Metadata")
vegdata <- read_xlsx("~/GitHub/BeccaTransects/data/VegData_raw.xlsx", sheet = "Data")
```

# 1. creating the main data frame  **Need to replace SaDe_new and SaPa_new with just SaDe and SaPa column names and then rerun plots before publishing!!:
```{r}
VegDataNew <- vegdata %>% 
  mutate_at(c('MeIn', 'DiGr', 'RaSa', 'CaMa', 'SaSo', 'Lolium', 'FrSa', 'Fabiaceae','Vicia', 'Sonc', 'Xant', 'LeTr', 'Ranunculus','Taraxacum', 'Coytote_Bush','BoMa', 'DiSp'), as.numeric) %>% 
  filter(!(BaGr_Type_1 == '7' | BaGr_Type_2 == '7')) %>% #remove rows where BaGr is coded as log
  mutate_if(is.numeric, list(~replace_na(., 0))) %>%
  select(-BaGr_Type_1, -BaGr_Type_2, -Coytote_Bush, -Wrack, -Tot_Num_Species, -Shoot_Density, -Length_centimeters, -Patch_Type) %>% 
  mutate(BaGr = rowSums(across(BaGr_1:BaGr_2), na.rm = T)) %>% 
  select(-BaGr_1, -BaGr_2) %>% 
  mutate(Total_Veg = 25-BaGr) %>%  #Total_veg is simply what is remaining after subtracting the bare ground
  mutate(across(SpFo:Total_Veg, ~ .x*4)) %>% 
  mutate(Date = paste0(Year, "-", Month)) %>% 
  mutate(Total_Veg2 = rowSums(across(SpFo:LeTr))) %>% 
  filter(!Total_Veg2 == 0) #get rid of all rows that equal 0

VegDataMerged <- full_join(VegDataNew, vegmetadata, by = "Transect_ID") %>% 
  select(-c(...9:...12)) %>% 
  mutate(SaSp = SaPa + SaDe) %>% 
  mutate(SaPa_new = SaPa + ((SaPa/SaSp)*PicklePups)) %>% 
  mutate(SaDe_new = SaDe + ((SaDe/SaSp)*PicklePups)) %>% 
  mutate(Grass = GrGR + LeTr + MixedSeedlings) %>% 
  mutate(across(everything(), ~replace(.x, is.nan(.x), 0)))
 



```

# #################################################################################################################### #

#this is for June of 2022! 

#2. Now I need to seperate my data into two df's for EACH ZONE -- one for identifiers and one for abundance data only.
# *We removed the row where C6-L174 is equal to quadrat 3 and zone U since it was a big outlier.
```{r}
VegDataID_U <- VegDataMerged[VegDataMerged$Date == "2022-June" & VegDataMerged$Zone == "U" &
                             !(VegDataMerged$Transect_ID == "C6-L174" & VegDataMerged$Quadrat == 3 & VegDataMerged$Zone == "U"),
                           c('Transect_ID', 'Distance_meters', 'Quadrat', 'Zone', 'Date', 'Shoreline_End', 'Log_Presence')]



VegDataSpecies_U <- VegDataMerged[(VegDataMerged$Date == "2022-June" & VegDataMerged$Zone == "U" &
                                  !(VegDataMerged$Transect_ID == "C6-L174" & 
                                      VegDataMerged$Quadrat == 3)), 
                                c(7:37, 50:52)] %>% 
  select(-c('GrGR','LeTr','MixedSeedlings','SaPa','SaDe','PicklePups')) #'CaMa', 'Lolium', 'RaSa', 'FrSa', 'Fabiaceae', 'Ranunculus', 'Taraxacum'))



#Transform the veg communities using the hellinger transformation.This transformation is useful for abundance data when you may have a few really abundant species and many rare species.It transforms the abundances of the really abundant and few rare species don't have less of an influence on the ordination. hellinger: square root of method = "total" (Legendre & Gallagher 2001).

VegDataSpecies.hel_U <- decostand(VegDataSpecies_U, method = "hellinger")


```


```{r}
VegDataID_M <- VegDataMerged[VegDataMerged$Date == "2022-June" & VegDataMerged$Zone == "M",
                             #!(VegDataMerged$Transect_ID == "C6-L174" & VegDataMerged$Quadrat == 3 & VegDataMerged$Zone == "U"),
                           c('Transect_ID', 'Distance_meters', 'Quadrat', 'Zone', 'Date', 'Shoreline_End', 'Log_Presence')]


VegDataSpecies_M <- VegDataMerged[VegDataMerged$Date == "2022-June" & VegDataMerged$Zone == "M",
                                  #!(VegDataMerged$Transect_ID == "C6-L174" & 
                                      #VegDataMerged$Quadrat == 3)),
                                c(7:37, 50:52)] %>% 
  select(-c('GrGR','LeTr','MixedSeedlings','SaPa','SaDe','PicklePups'))

#Transform the veg communities using the hellinger transformation.This transformation is useful for abundance data when you may have a few really abundant species and many rare species.It transforms the abundances of the really abundant and few rare species don't have less of an influence on the ordination. hellinger: square root of method = "total" (Legendre & Gallagher 2001).

VegDataSpecies.hel_M <- decostand(VegDataSpecies_M, method = "hellinger")
```

```{r}
VegDataID_L <- VegDataMerged[VegDataMerged$Date == "2022-June" & VegDataMerged$Zone == "L",
                             #!(VegDataMerged$Transect_ID == "C6-L174" & VegDataMerged$Quadrat == 3 & VegDataMerged$Zone == "U"),
                           c('Transect_ID', 'Distance_meters', 'Quadrat', 'Zone', 'Date', 'Shoreline_End', 'Log_Presence')]


VegDataSpecies_L <- VegDataMerged[VegDataMerged$Date == "2022-June" & VegDataMerged$Zone == "L",
                                  #!(VegDataMerged$Transect_ID == "C6-L174" & 
                                      #VegDataMerged$Quadrat == 3)),
                                c(7:37, 50:52)] %>% 
  select(-c('GrGR','LeTr','MixedSeedlings','SaPa','SaDe','PicklePups'))

#Transform the veg communities using the hellinger transformation.This transformation is useful for abundance data when you may have a few really abundant species and many rare species.It transforms the abundances of the really abundant and few rare species don't have less of an influence on the ordination. hellinger: square root of method = "total" (Legendre & Gallagher 2001).

VegDataSpecies.hel_L <- decostand(VegDataSpecies_L, method = "hellinger")
```

# ################################################################################################################## #
#3. Running the NMDS:


# Per Margot: Run the nMDS. This is where I previously had an extra step, which I removed. You can play with the value of k, which is the number of dimensions. #2 is best but if you increase dimensions to 3 you decrease stress, which might be necessary for high values.
#Stress: 0-0.1 = excellent fit, 0.1 – 0.2 = good fit, 0.2 – 0.3 = very mediocre fit, 0.3+ = trash
#The default is trymax = 20. This is often not enough to achieve convergence, so I often increase this value
#I generally don’t increase it past 10,000 – anything more than that and your data is not going to work with the analysis, or you need to mess with k
# *** If they are all hashtagged out it is to preserve the last run of the MDS. Everything is all good.
```{r}
### UPPER NMDS ###
set.seed(19840217)
#VegDataSpeciesNMDS_U <- metaMDS(VegDataSpecies_U, k = 2, trymax = 20)
#VegDataSpeciesNMDS_U 

#VegDataSpeciesNMDS_U  <- metaMDS(VegDataSpecies_U, previous.best = VegDataSpeciesNMDS_U)




### Hellinger Method ###

nmds_U <- metaMDS(VegDataSpecies.hel_U, autotransform = F, k = 2, trymax = 1000) #autotransform = F so that it doesn't revert back to the default transformation *bc we are using the hellinger method.
nmds_U

save(nmds_U, file ="../output_data/nmds_U_June2022.Rds")

load("../output_data/nmds_U_June2022.Rds")
```

```{r}
### MIDDLE NMDS ###
set.seed(19840217)
#VegDataSpeciesNMDS_M <- metaMDS(VegDataSpecies_M, k = 2, trymax = 20)
#VegDataSpeciesNMDS_M 

#VegDataSpeciesNMDS_M  <- metaMDS(VegDataSpecies_M, previous.best = VegDataSpeciesNMDS_U)




## Hellinger Method ###

nmds_M <- metaMDS(VegDataSpecies.hel_M, autotransform = F, k = 2, trymax = 1000) #autotransform = F so that it doesn't revert back to the default transformation *bc we are using the hellinger method.
nmds_M

save(nmds_M, file ="../output_data/nmds_M_June2022.Rds")

load("../output_data/nmds_M_June2022.Rds")
```

```{r}
### LOWER NMDS ###
set.seed(19840217)
#VegDataSpeciesNMDS_L <- metaMDS(VegDataSpecies_L, k = 2, trymax = 20)
#VegDataSpeciesNMDS_L 

#VegDataSpeciesNMDS_L  <- metaMDS(VegDataSpecies_L, previous.best = VegDataSpeciesNMDS_U)




## Hellinger Method ###

nmds_L <- metaMDS(VegDataSpecies.hel_L, autotransform = F, k = 2, trymax = 1000) #autotransform = F so that it doesn't revert back to the default transformation *bc we are using the hellinger method.
nmds_L

save(nmds_L, file ="../output_data/nmds_L_June2022.Rds")

load("../output_data/nmds_L_June2022.Rds")
```




# ################################################################################################################################## #
#4. Plotting 


#Starting with Upper Zone:
```{r}
###  Plotting Hellinger method in base R ###
nmds_U

ordiplot(nmds_U)
ordiplot(nmds_U, type = "t") #puts species name on plot


# using ggplot with Hellinger Method ###

NMDS1 <- as.numeric(as.character(nmds_U$points[,1]))
NMDS2 <- as.numeric(as.character(nmds_U$points[,2]))
mds.coords1_U <- data.frame(VegDataID_U, NMDS1, NMDS2)

cbp <- c("#009E73", "#D55E00", "#CC79A7","#56B4E9", "#F0E442" , "#E69F00", "#000000", "#0072B2")


EWLNLplot <- ggplot() +
  geom_point(data = mds.coords1_U, aes(x=NMDS1, y=NMDS2, 
                                       color = Log_Presence,
                                       shape = Shoreline_End,
                                       alpha = 0.8)) +
  scale_color_manual(values = cbp) +   #override base ggplot colors with colorblind friendly colors
  #geom_segment(data = (mds.coords1_U),
  #             mapping = aes(x =0, y = 0, xend = NMDS1, yend = NMDS2),
  #             arrow = arrow(length = unit(0.015, "npc"),
  #                           type ="closed"),
  #             color = "darkgray",
  #            alpha = 0.5,
  #             size = 0.6) +
               labs(title = "Upper Zone nMDS",
                    subtitle = "stress = 0.10, bray") +
  geom_abline(intercept=0, slope=0, linetype = "dashed", size = 0.8, color = "gray")+
  geom_vline(aes(xintercept=0), linetype = "dashed", size = 0.8, color = "gray")+
  labs(title = "Upper Zone nMDS") +
  theme_bw() +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(color = "black"))

EWLNLplot



##experimenting with getting species names on ggplot using Hellinger method.  Has species names but not E v. W or L v. NL info:

fort <- fortify(nmds_U) #allow the data contained within objects created by vegan to be converted into a format suitable for use with ggplot directly.

ggplot() +
  geom_point(data = subset(fort, score == 'sites'),
             mapping = aes(x=NMDS1, y=NMDS2),
             color = "#D55E00",
             alpha = 0.5) +
  geom_segment(data = subset(fort, score == 'species'),
               mapping = aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2),
               arrow = arrow(length = unit(0.015, "npc"),
                             type ="closed"),
               color = "darkgray",
               size = 0.6) +
               labs(title = "Upper Zone nMDS",
                    subtitle = "stress = 0.0830904, bray") +
               theme_bw() +
  geom_text(data = subset(fort, score == 'species'),
            mapping = aes(label=label, x=NMDS1*1.1, y=NMDS2*1.1)) +
  geom_abline(intercept=0, slope=0, linetype = "dashed", size = 0.8, color = "gray")+
  geom_vline(aes(xintercept=0), linetype = "dashed", size = 0.8, color = "gray")+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(color = "black"))

# Interpretation: The points, or "sites" that were sampled by the nmds, that are > 0 on the x-axis were characterized by higher abundances of those species
# that have positive values along the x-axis. Conversely,the sites with negative values have low abundances of the species listed < 0.





#Making a two panel plot:

#points only.  NO arrows:
p1 <- ggplot() +
  geom_point(data = subset(fort, score == 'sites'),
             mapping = aes(x=NMDS1, y=NMDS2),
             color = "#D55E00",
             alpha = 0.5) +
  geom_segment(data = subset(fort, score == 'species'),
               mapping = aes(x =0, y = 0, xend = NMDS1, yend = NMDS2),
               arrow = arrow(length = unit(0.015, "npc"),
                             type ="closed"),
               color = "darkgray",
               size = 0,     #set size and alpha = 0 so that the arrows are scaled the same as p2 but they don't actually show up.
               alpha = 0) +
  geom_text(data = subset(fort, score == 'species'),
            mapping = aes(label=label, x=NMDS1*1.1, y=NMDS2*1.1),
            alpha = 0) +     #also set alpha = o here, too.
  scale_y_continuous(limits=c(-2, 2)) +  #make the y=axis the same scale as the y-axis on plots p3 and p2
  labs(title = "Upper Zone nMDS",
       subtitle = "stress = 0.0830904, bray") +
  theme_bw() +
  geom_abline(intercept=0, slope=0, linetype = "dashed", size = 0.8, color = "gray")+
  geom_vline(aes(xintercept=0), linetype = "dashed", size = 0.8, color = "gray")+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(face = "bold", size = 16),
        axis.title.x = element_blank(),    #remove the x-axis title
        axis.line = element_line(color = "black")) 

p1   #points, but no species labels or arrows


p2 <- ggplot() +
  geom_point(data = subset(fort, score == 'sites'),
             mapping = aes(x=NMDS1, y=NMDS2),
             color = "#D55E00",
             alpha = 0) +     #set alpha = 0 here so that the points are fully transparent.
  geom_segment(data = subset(fort, score == 'species'),
               mapping = aes(x =0, y = 0, xend = NMDS1, yend = NMDS2),
               arrow = arrow(length = unit(0.015, "npc"),
                             type ="closed"),
               color = "darkgray",
               size = 0.6) +
               theme_bw() +
  geom_text(data = subset(fort, score == 'species'),
            mapping = aes(label=label, x=NMDS1*1.1, y=NMDS2*1.1)) +
  scale_y_continuous(limits=c(-2, 2)) +  #make the y=axis the same scale as the y-axis on plots p3 and p1
  labs(title = "Upper Zone nMDS - June 2022",
       subtitle = "stress = 0.0830904, bray") +
  geom_abline(intercept=0, slope=0, linetype = "dashed", size = 0.8, color = "gray")+
  geom_vline(aes(xintercept=0), linetype = "dashed", size = 0.8, color = "gray")+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(face = "bold", size = 16),
        axis.title.x = element_blank(),    #remove the x-axis title
        axis.line = element_line(color = "black"))
p2   #species and arrow, but no points



#p3 below uses the data from EWLNLplot above with a few modifications
p3 <- ggplot() +
  geom_point(data = mds.coords1_U, 
             aes(x=NMDS1, y=NMDS2, 
                color = Log_Presence,
                shape = Shoreline_End),
             alpha = 0.5,
             size = 4) +    #keep this detail outside of the aes() if you don't want it to show up in the legend
  scale_color_manual(values = cbp) +   #override base ggplot colors with colorblind friendly colors
  scale_y_continuous(limits=c(-2, 2)) +  #make the y=axis the same scale as the y-axis on plots p1 and p2
  #labs(title = "Upper Zone nMDS",
  #    subtitle = "stress = 0.100779, bray") +
  geom_abline(intercept=0, slope=0, linetype = "dashed", size = 0.8, color = "gray")+
  geom_vline(aes(xintercept=0), linetype = "dashed", size = 0.8, color = "gray")+
  
  theme_bw() +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(color = "black"),
        legend.position = "bottom", # move legend to bottom
        legend.title = element_text(face = "bold", size=10),
        legend.text=element_text(size=8)) 

p3


#Now plot:
ggarrange(p2, p3, ncol=1,     #left out p1 bc it is redundant with p3
          common.legend = TRUE,
          legend = "bottom")
```
#Test for differences in the veg community composition across shoreline end and log v no log.
```{r}
#First, need to remove these previous columns that were combined back in the beginning and remove the weirdo in quad 3 for C6-L174:
VegDataMerged_mod_U <- VegDataMerged[(VegDataMerged$Date == "2023-August" & VegDataMerged$Zone == "U" &
                                  !(VegDataMerged$Transect_ID == "C6-L174" & 
                                      VegDataMerged$Quadrat == 3)), ] %>% 
  select(-c('GrGR','LeTr','MixedSeedlings','SaPa','SaDe','PicklePups'))

### Not sure if I can run this or not!!!
adonis2(Total_Veg ~ Shoreline_End, data=VegDataMerged_mod_U)


```


#Middle Zone
```{r}
###  Plotting Hellinger method in base R ###

ordiplot(nmds_M)
ordiplot(nmds_M, type = "t") #puts species name on plot

nmds_M

# using ggplot with Hellinger Method ###

NMDS1 <- as.numeric(as.character(nmds_M$points[,1]))
NMDS2 <- as.numeric(as.character(nmds_M$points[,2]))
mds.coords1_M <- data.frame(VegDataID_M, NMDS1, NMDS2)

cbp <- c("#009E73", "#D55E00", "#CC79A7","#56B4E9", "#F0E442" , "#E69F00", "#000000", "#0072B2")


EWLNL_M_plot <- ggplot() +
  geom_point(data = mds.coords1_M, aes(x=NMDS1, y=NMDS2, 
                                       color = Log_Presence,
                                       shape = Shoreline_End,
                                       alpha = 0.8)) +
  scale_color_manual(values = cbp) +   #override base ggplot colors with colorblind friendly colors
  #geom_segment(data = (mds.coords1_U),
  #             mapping = aes(x =0, y = 0, xend = NMDS1, yend = NMDS2),
  #             arrow = arrow(length = unit(0.015, "npc"),
  #                           type ="closed"),
  #             color = "darkgray",
  #            alpha = 0.5,
  #             size = 0.6) +
               labs(title = "Middle Zone nMDS",
                    subtitle = "stress = 0.06263913, bray") +
  geom_abline(intercept=0, slope=0, linetype = "dashed", size = 0.8, color = "gray")+
  geom_vline(aes(xintercept=0), linetype = "dashed", size = 0.8, color = "gray")+
  labs(title = "Middle Zone nMDS") +
  theme_bw() +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(color = "black"))

EWLNL_M_plot



##experimenting with getting species names on ggplot using Hellinger method.  Has species names but not E v. W or L v. NL info:

fort <- fortify(nmds_M) #allow the data contained within objects created by vegan to be converted into a format suitable for use with ggplot directly. If it says "data` must be a <data.frame>, or an object coercible by `fortify()`, not a <metaMDS/monoMDS> object," reload ggvegan package.

ggplot() +
  geom_point(data = subset(fort, score == 'sites'),
             mapping = aes(x=NMDS1, y=NMDS2),
             color = "#D55E00",
             alpha = 0.5) +
  geom_segment(data = subset(fort, score == 'species'),
               mapping = aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2),
               arrow = arrow(length = unit(0.015, "npc"),
                             type ="closed"),
               color = "darkgray",
               size = 0.6) +
               labs(title = "Middle Zone nMDS",
                    subtitle = "stress = 0.06263913, bray") +
               theme_bw() +
  geom_text(data = subset(fort, score == 'species'),
            mapping = aes(label=label, x=NMDS1*1.1, y=NMDS2*1.1)) +
  geom_abline(intercept=0, slope=0, linetype = "dashed", size = 0.8, color = "gray")+
  geom_vline(aes(xintercept=0), linetype = "dashed", size = 0.8, color = "gray")+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(color = "black"))

# Interpretation: The points, or "sites" that were sampled by the nmds, that are > 0 on the x-axis were characterized by higher abundances of those species
# that have positive values along the x-axis. Conversely,the sites with negative values have low abundances of the species listed < 0.





#Making a two panel plot:

#points only.  NO arrows:
p1 <- ggplot() +
  geom_point(data = subset(fort, score == 'sites'),
             mapping = aes(x=NMDS1, y=NMDS2),
             color = "#D55E00",
             alpha = 0.5) +
  geom_segment(data = subset(fort, score == 'species'),
               mapping = aes(x =0, y = 0, xend = NMDS1, yend = NMDS2),
               arrow = arrow(length = unit(0.015, "npc"),
                             type ="closed"),
               color = "darkgray",
               size = 0,     #set size and alpha = 0 so that the arrows are scaled the same as p2 but they don't actually show up.
               alpha = 0) +
  geom_text(data = subset(fort, score == 'species'),
            mapping = aes(label=label, x=NMDS1*1.1, y=NMDS2*1.1),
            alpha = 0) +     #also set alpha = o here, too.
  scale_y_continuous(limits=c(-2, 2)) +  #make the y=axis the same scale as the y-axis on plots p3 and p2
  labs(title = "Middle Zone nMDS",
       subtitle = "stress = 0.06263913, bray") +
  theme_bw() +
  geom_abline(intercept=0, slope=0, linetype = "dashed", size = 0.8, color = "gray")+
  geom_vline(aes(xintercept=0), linetype = "dashed", size = 0.8, color = "gray")+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(face = "bold", size = 16),
        axis.title.x = element_blank(),    #remove the x-axis title
        axis.line = element_line(color = "black")) 

p1   #points, but no species labels or arrows


p2 <- ggplot() +
  geom_abline(intercept=0, slope=0, linetype = "dashed", size = 0.8, color = "gray")+   #putting these lines under the species labels
  geom_vline(aes(xintercept=0), linetype = "dashed", size = 0.8, color = "gray")+
  geom_point(data = subset(fort, score == 'sites'),
             mapping = aes(x=NMDS1, y=NMDS2),
             color = "#D55E00",
             alpha = 0) +     #set alpha = 0 here so that the points are fully transparent.
  geom_segment(data = subset(fort, score == 'species'),
               mapping = aes(x =0, y = 0, xend = NMDS1, yend = NMDS2),
               arrow = arrow(length = unit(0.015, "npc"),
                             type ="closed"),
               color = "darkgray",
               size = 0.6) +
               theme_bw() +
  geom_text(data = subset(fort, score == 'species'),
            mapping = aes(label=label, x=NMDS1*1.1, y=NMDS2*1.1)) +
  scale_y_continuous(limits=c(-2, 2)) +  #make the y=axis the same scale as the y-axis on plots p3 and p1
  scale_x_continuous(limits = c(-1, 1)) +
  labs(title = "Middle Zone nMDS - June 2022",
       subtitle = "stress = 0.06263913, bray") +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(face = "bold", size = 16),
        axis.title.x = element_blank(),    #remove the x-axis title
        axis.line = element_line(color = "black"))
p2   #species and arrow, but no points



#p3 below uses the data from EWLNLplot above with a few modifications
p3 <- ggplot() +
  geom_point(data = mds.coords1_M, 
             aes(x=NMDS1, y=NMDS2, 
                color = Log_Presence,
                shape = Shoreline_End),
             alpha = 0.5,
             size = 4) +          #keep this detail outside of the aes() if you don't want it to show up in the legend
  scale_color_manual(values = cbp) +   #override base ggplot colors with colorblind friendly colors
  scale_y_continuous(limits=c(-2, 2)) +  #make the y=axis the same scale as the y-axis on plots p1 and p2
  scale_x_continuous(limits = c(-1, 1)) +
  #scale_x_(limits = -1, 1) +
  #labs(title = "Upper Zone nMDS",
  #    subtitle = "stress = 0.100779, bray") +
  geom_abline(intercept=0, slope=0, linetype = "dashed", size = 0.8, color = "gray")+
  geom_vline(aes(xintercept=0), linetype = "dashed", size = 0.8, color = "gray")+
  
  theme_bw() +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(color = "black"),
        legend.position = "bottom", # move legend to bottom
        legend.title = element_text(face = "bold", size=10),
        legend.text=element_text(size=8)) 

p3


#Now plot:
ggarrange(p2, p3, ncol=1,     #left out p1 bc it is redundant with p3
          common.legend = TRUE,
          legend = "bottom")
```


#Lower Zone:
```{r}
###  Plotting Hellinger method in base R ###
nmds_L

ordiplot(nmds_L)
ordiplot(nmds_L, type = "t") #puts species name on plot



# using ggplot with Hellinger Method ###

NMDS1 <- as.numeric(as.character(nmds_L$points[,1]))
NMDS2 <- as.numeric(as.character(nmds_L$points[,2]))
mds.coords1_L <- data.frame(VegDataID_L, NMDS1, NMDS2)

cbp <- c("#009E73", "#D55E00", "#CC79A7","#56B4E9", "#F0E442" , "#E69F00", "#000000", "#0072B2")


EWLNL_L_plot <- ggplot() +
  geom_point(data = mds.coords1_L, aes(x=NMDS1, y=NMDS2, 
                                       color = Log_Presence,
                                       shape = Shoreline_End,
                                       alpha = 0.8)) +
  scale_color_manual(values = cbp) +   #override base ggplot colors with colorblind friendly colors
  #geom_segment(data = (mds.coords1_U),
  #             mapping = aes(x =0, y = 0, xend = NMDS1, yend = NMDS2),
  #             arrow = arrow(length = unit(0.015, "npc"),
  #                           type ="closed"),
  #             color = "darkgray",
  #            alpha = 0.5,
  #             size = 0.6) +
  labs(title = "Lower Zone nMDS",
  subtitle = "stress = 0.05146037 , bray") +
  geom_abline(intercept=0, slope=0, linetype = "dashed", size = 0.8, color = "gray")+
  geom_vline(aes(xintercept=0), linetype = "dashed", size = 0.8, color = "gray")+
  
  theme_bw() +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(color = "black"))

EWLNL_L_plot



##experimenting with getting species names on ggplot using Hellinger method.  Has species names but not E v. W or L v. NL info:

fort <- fortify(nmds_L) #allow the data contained within objects created by vegan to be converted into a format suitable for use with ggplot directly. If it says "data` must be a <data.frame>, or an object coercible by `fortify()`, not a <metaMDS/monoMDS> object," reload ggvegan package.

ggplot() +
  geom_point(data = subset(fort, score == 'sites'),
             mapping = aes(x=NMDS1, y=NMDS2),
             color = "#D55E00",
             alpha = 0.5) +
  geom_segment(data = subset(fort, score == 'species'),
               mapping = aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2),
               arrow = arrow(length = unit(0.015, "npc"),
                             type ="closed"),
               color = "darkgray",
               size = 0.6) +
               labs(title = "Lower Zone nMDS",
                    subtitle = "stress = 0.05146037, bray") +
               theme_bw() +
  geom_text(data = subset(fort, score == 'species'),
            mapping = aes(label=label, x=NMDS1*1.1, y=NMDS2*1.1)) +
  geom_abline(intercept=0, slope=0, linetype = "dashed", size = 0.8, color = "gray")+
  geom_vline(aes(xintercept=0), linetype = "dashed", size = 0.8, color = "gray")+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(color = "black"))

# Interpretation: The points, or "sites" that were sampled by the nmds, that are > 0 on the x-axis were characterized by higher abundances of those species
# that have positive values along the x-axis. Conversely,the sites with negative values have low abundances of the species listed < 0.





#Making a two panel plot:

#points only.  NO arrows:
p1 <- ggplot() +
  geom_point(data = subset(fort, score == 'sites'),
             mapping = aes(x=NMDS1, y=NMDS2),
             color = "#D55E00",
             alpha = 0.5) +
  geom_segment(data = subset(fort, score == 'species'),
               mapping = aes(x =0, y = 0, xend = NMDS1, yend = NMDS2),
               arrow = arrow(length = unit(0.015, "npc"),
                             type ="closed"),
               color = "darkgray",
               size = 0,     #set size and alpha = 0 so that the arrows are scaled the same as p2 but they don't actually show up.
               alpha = 0) +
  geom_text(data = subset(fort, score == 'species'),
            mapping = aes(label=label, x=NMDS1*1.1, y=NMDS2*1.1),
            alpha = 0) +     #also set alpha = o here, too.
  scale_y_continuous(limits=c(-2, 2)) +  #make the y=axis the same scale as the y-axis on plots p3 and p2
  labs(title = "Lower Zone nMDS",
       subtitle = "stress = 0.05146037, bray") +
  theme_bw() +
  geom_abline(intercept=0, slope=0, linetype = "dashed", size = 0.8, color = "gray")+
  geom_vline(aes(xintercept=0), linetype = "dashed", size = 0.8, color = "gray")+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(face = "bold", size = 16),
        axis.title.x = element_blank(),    #remove the x-axis title
        axis.line = element_line(color = "black")) 

p1   #points, but no species labels or arrows


p2 <- ggplot() +
  geom_abline(intercept=0, slope=0, linetype = "dashed", size = 0.8, color = "gray")+   #putting these lines under the species labels
  geom_vline(aes(xintercept=0), linetype = "dashed", size = 0.8, color = "gray")+
  geom_point(data = subset(fort, score == 'sites'),
             mapping = aes(x=NMDS1, y=NMDS2),
             color = "#D55E00",
             alpha = 0) +     #set alpha = 0 here so that the points are fully transparent.
  geom_segment(data = subset(fort, score == 'species'),
               mapping = aes(x =0, y = 0, xend = NMDS1, yend = NMDS2),
               arrow = arrow(length = unit(0.015, "npc"),
                             type ="closed"),
               color = "darkgray",
               size = 0.6) +
               theme_bw() +
  geom_text(data = subset(fort, score == 'species'),
            mapping = aes(label=label, x=NMDS1*1.1, y=NMDS2*1.1)) +
  scale_y_continuous(limits=c(-2, 2)) +  #make the y=axis the same scale as the y-axis on plots p3 and p1
  scale_x_continuous(limits = c(-1, 1)) +
  labs(title = "Lower Zone nMDS - June 2022",
       subtitle = "stress = 0.05146037, bray") +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        plot.title = element_text(face = "bold", size = 16),
        axis.title.x = element_blank(),    #remove the x-axis title
        axis.line = element_line(color = "black"))
p2   #species and arrow, but no points



#p3 below uses the data from EWLNLplot above with a few modifications
p3 <- ggplot() +
  geom_point(data = mds.coords1_L, 
             aes(x=NMDS1, y=NMDS2, 
                color = Log_Presence,
                shape = Shoreline_End),
             alpha = 0.5,
             size = 4) +          #keep this detail outside of the aes() if you don't want it to show up in the legend
  scale_color_manual(values = cbp) +   #override base ggplot colors with colorblind friendly colors
  scale_y_continuous(limits=c(-2, 2)) +  #make the y=axis the same scale as the y-axis on plots p1 and p2
  scale_x_continuous(limits = c(-1, 1)) +
  #scale_x_(limits = -1, 1) +
  #labs(title = "Upper Zone nMDS",
  #    subtitle = "stress = 0.100779, bray") +
  geom_abline(intercept=0, slope=0, linetype = "dashed", size = 0.8, color = "gray")+
  geom_vline(aes(xintercept=0), linetype = "dashed", size = 0.8, color = "gray")+
  
  theme_bw() +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(color = "black"),
        legend.position = "bottom", # move legend to bottom
        legend.title = element_text(face = "bold", size=10),
        legend.text=element_text(size=8)) 

p3


#Now plot:
ggarrange(p2, p3, ncol=1,     #left out p1 bc it is redundant with p3
          common.legend = TRUE,
          legend = "bottom")
```